\section{Abstract}
The goal of this project is to integrate the Dafny programming language into Visual Studio Code. Emphasis is put into researching how Dafny programmers can be best supported during their work and how writing code can be made more productive. \newline

Since Dafny offers built-in specification constructs, novel work is to provide tooling that makes working with them easier. The most beneficial feature would be to implement a generic, context aware proof obligation generator that continuously suggests specification constructs to the programmer. This approach was eventually discarded because it was deemed to hard to make it work with the proof engine that Dafny uses. Instead, situations were identified that arise often during programming and specific aid with specification constructs was implemented for those situations. Another helpful Dafny specific features is the displaying of counter examples where the code written does not satisfy the corresponding specification constructs, allowing quick discovery of edge cases and refinement of specification constructs. \newline

Next to language specific features, standard IDE mechanisms allow for great improvement regarding productivity. It was therefor deemed paramount that the project implements the most common features such as go to definition, renaming refactorings and find references. This was achieved using the standard interfaces that Visual Studio Code provides for these use cases, allowing programmers accessing these features in a well established way. \newline

Regarding the fact that Dafny is still a very young language, it is an important concern that new users can get started quickly, so that the user base continuous to grow. To support this, automatic installation of the plugin on all major platforms was implemented. Next to the plugin itself, the installation also resolves all dependencies such as the Dafny compiler itself. To further maximize portability, the plugin implements the language server protocol. This allows for writing IDE agnostic language analysis platforms, making the plugin not only usable in Visual Studio Code, but also integrable into other IDEs such as Emacs with only minor adjustments. \newline

This project concluded with the implementation of a production ready integration of Dafny into Visual Studio Code. The application of continuous integration allowed for a user base of about 300 people at the end of this writing, proofing that the plugin is robust and works across multiple environments. Dafny programmers are supported in their coding not only with standard IDE mechanisms, but also Dafny specific features. This not only makes the experience of programming more productive, but lays the foundation of a contentiously growing Dafny community. \newline
