\section{Use Cases}
\subsection{Use Case Diagramm}
\subsection{Aktoren und Stakeholder}
\subsection{Descriptions (brief)}
\subsubsection{UC1: Stable working Release on Windows}
A User can install the Dafny Plugin on his .Net environment running on Windows 10.
\subsubsection{UC2: Stable working Release on Linux}
A User can install the Dafny Plugin on his mono environment running on a common Linux distribution such as Ubuntu 16.04.
\subsubsection{UC3: Stable working Release on OSX}
A User can install the Dafny Plugin on his mono environment running on OSX.
\subsubsection{UC4: Easy installation of Dafny plugin}
A User can simply install the Dafny plugin by running an automatic installer which sets all path variables and makes additional needed environment adjustments.
\subsubsection{UC5: Syntax Highlighting}
The system automatically does syntax highlighting while the user writes a Dafny (dy.) file.
\subsubsection{UC6: Compilation of Dafny best practices and reporting of these rules}
The plugin can be configured with a simple json-like config file which describes common best practices for Dafny. The Plugin reports violations of these rules via the standard Visual Studio Code notification mechanisms.
\subsubsection{UC7: Automatic generation of contracts}
The plugin can be configured with a simple DSL-File to recognize certain semantics, which could benefit from the setting of Pre- and Postconditions. The plugin offers to add these in form of a refactoring via the common Visual Studio Code mechanisms.
\subsubsection{UC8: Autocompletion for identifiers}
The plugin offers autocompletion of Dafny code while the user types via the standard Visual Studio Code mechanisms.
\subsection{Descriptions (fully dressed)}