\section{Dafny VSCode}
Github Repository: \href{https://github.com/FunctionalCorrectness/dafny-vscode}{https://github.com/FunctionalCorrectness/dafny-vscode}

\subsection{Overview}
\todo{Maybe add a class diagram}

\subsection{Visual Studio Code Plugin}
\subsubsection{Structure}
When a Visual Studio Code plugin is implemented as a language server, the plugin consists of two parts. The first one is an IDE-agnostic language server, which is further explained in \ref{langServer}. The second part is the actual plugin which interacts with Visual Studio Code and relies on the language server to provide functionality. \newline
The actual plugin will further on be referenced as the client part of the plugin. The most important part of the client part is a file called package.json. It informs the IDE about the following points:
\paragraph{Activation}
Details for what the plugin is designed and when it is activated. Normally, plugins stay dormant until a use case is triggered for which they are needed. For this project, this is the case whenever a Dafny file is opened in Visual Studio Code.

\paragraph{Commands}
Details which commands this plugin exposes. Next to standard callbacks such as when a file is first opened or saved, additional custom commands can be registered here. This projects exposes some custom commands such as to install the Dafny server or check the version of the server currently installed. This also includes key bindings for the registered commands.

\paragraph{Configuration}
Is used to configure a plugin. For this project this for instance entails the path to the Dafny executable or if it is running on a .NET or mono environment. It also sets Visual Studio Code specific configuration, such as how long of a delay should follow typing before the file is reevaluated.

\paragraph{Dependencies}
Details all the dependencies in the form of node modules that the plugin has.

\paragraph{Meta information}
Additionally, this section details meta information about the plugin such as who developed it, which version it has and so on.


\todo{Further explanation about the client side of the plugin}


\input{sections/languageserver}

\subsection{Extension points}

\paragraph{Status bar}

\paragraph{Refactoring}

\paragraph{CodeLens}

\paragraph{Go to definition}

\paragraph{Debugger}
\todo{Boogie Verification Debugger: Why BVD can't be used directly: C sharp UI, }

\subsection{Communication}

\subsection{Syntax highlighting}

\subsection{Snippets}

\subsection{Automatic installation}
\subsubsection{Windows}

\subsubsection{Ubuntu}

\subsubsection{OSX}


\subsection{DafnyServer}
Github Repository: \href{https://github.com/FunctionalCorrectness/dafny-microsoft}{https://github.com/FunctionalCorrectness/dafny-microsoft}

The DafnyServer is a simple console application which allows proofing Dafny source files. To verify documents, they are sent over the standard input. Results are obtained from the standard output. The verification task needs to be in JSON format \todo{add reference to source.cs} and is sent base64 encoded. By default, the server only supports the verbs verify, quit and selftest. Verbs are sent first, followed by a newline \textbackslash{n}. They may be proceeded by a payload and an end string \textbf{[[DAFNY-CLIENT: EOM]]}. \newline 
Verb explanation: Verify needs a verification task and returns if all proofs holds, quit stops the server and selftest execute some simple verification. \newline

\textbf{Example verification task}
\begin{lstlisting}[language=json,firstnumber=1]
{
	args:[],
	filename:"c:\Users\Markus\Desktop\dafny\test1.dfy",
	source:"method Main() {	assert 1 < 3; }",
	sourceIsFile:false
}

\end{lstlisting}

\subsubsection{symbols}
To support refactoring in the Dafny Visual Studio Code plugin, symbol information was needed. All fields, methods and classes inside a file along with their information about position, reference and usage have to be accessible. To support this, the DafnyServer was extended. A new verb "symbols" was introduced. This collects various information about the symbol table of the input file and returns it as JSON. 
\newline\newline
\textbf{Request: }

\begin{lstlisting}[language=dafny]
class BankAccountUnsafe {
  var balance: int;
  constructor() modifies this { 
    balance := 10;
  }
  
  method withdraw(amount: int) 
    modifies this
    requires amount >= 0
  {   
    balance := balance - amount; 
  } 
}   

method test() { 
  var a := new BankAccountUnsafe(); 
  a.withdraw(9);  
}   
\end{lstlisting}

\textbf{Result: }
\begin{lstlisting}[language=json,firstnumber=1]
[ 
....
{
  "Call" : null,
  "Column" : 3,
  "EndColumn" : null,
  "EndLine" : null,
  "EndPosition" : null,
  "Ensures" : [],
  "Line" : 3,
  "Module" : "_module",
  "Name" : "_ctor",
  "ParentClass" : "BankAccountUnsafe",
  "Position" : 49,
  "References" : [{
    "Column" : 12,
    "Line" : 15,
    "MethodName" : "test",
    "Position" : 265,
    "ReferencedName" : "_ctor"
  }],
  "Requires" : [],
  "SymbolType" : "Method"
}, {
  "Call" : null,
  "Column" : 9,
  "EndColumn" : null,
  "EndLine" : null,
  "EndPosition" : null,
  "Ensures" : [],
  "Line" : 6,
  "Module" : "_module",
  "Name" : "withdraw",
  "ParentClass" : "BankAccountUnsafe",
  "Position" : 114,
  "References" : [{
    "Column" : 5,
    "Line" : 16,
    "MethodName" : "test",
    "Position" : 296,
    "ReferencedName" : "withdraw"
  }],
  "Requires" : ["amount >= 0"],
  "SymbolType" : "Method"
}, {
  "Call" : null,
  "Column" : 6,
  "EndColumn" : null,
  "EndLine" : null,
  "EndPosition" : null,
  "Ensures" : null,
  "Line" : 2,
  "Module" : "_module",
  "Name" : "balance",
  "ParentClass" : "BankAccountUnsafe",
  "Position" : 32,
  "References" : [{
    "Column" : 5,
    "Line" : 4,
    "MethodName" : "balance",
    "Position" : 85,
    "ReferencedName" : "balance"
  }, {
    "Column" : 3,
    "Line" : 10,
    "MethodName" : "balance",
    "Position" : 190,
    "ReferencedName" : "balance"
  }, {
    "Column" : 14,
    "Line" : 10,
    "MethodName" : "balance",
    "Position" : 201,
    "ReferencedName" : "balance"
  }],
  "Requires" : null,
  "SymbolType" : "Field"
}
.... 
]
\end{lstlisting}



\subsubsection{versioncheck}
\todo{Remove}
To support automatic updates of the Dafny environment in a general approach, it was decided to implement this features directly into the DafnyServer. To make it as general as possible, the check is directly performed on release page of the Dafny project. The GitHub API allows to check this in just one call. Additionally, a semantic versioning library\cite{semver} was used to compare the current and the version from GitHub. Based on that, it either checks if an update is available or if the latest version is installed. If there is an update available, the URL to download the package can also be read out of the returned JSON 
\newline
\textbf{Response from GitHub}
\begin{lstlisting}[language=json,firstnumber=1]
[{
  "name" : "v1.9.12",
  "assets" : [{
    "name" : "dafny-1.9.12-x64-osx-10.11.6.zip",
    "browser_download_url" : "https://github.com/FunctionalCorrectness/dafny-microsoft/releases/download/v1.9.12/dafny-1.9.12-x64-osx-10.11.6.zip"
  }, {
    "name" : "dafny-1.9.12-x64-ubuntu-14.04.zip",
    "browser_download_url" : "https://github.com/FunctionalCorrectness/dafny-microsoft/releases/download/v1.9.12/dafny-1.9.12-x64-ubuntu-14.04.zip"
  }, {
    "name" : "dafny-1.9.12-x64-win.zip",
    "browser_download_url" : "https://github.com/FunctionalCorrectness/dafny-microsoft/releases/download/v1.9.12/dafny-1.9.12-x64-win.zip"
  }]
}]
\end{lstlisting}

Currently the check is performed against the URL:  \href{https://api.github.com/repos/FunctionalCorrectness/dafny-microsoft/releases}{https://api.github.com/repos/FunctionalCorrectness/dafny-microsoft/releases} which has to be changed as soon as the pull request is performed. 


\subsubsection{version}
The command returns the version of the DafnyServer.  


\subsection{Dafny}

\subsubsection{Unicode}
\todo{Remove}
Dafny would support the following Unicode characters: 

\rowcolors{2}{gray!25}{white}
\begin{longtable}[H]
	{l|p{0.22\textwidth}| p{0.22\textwidth} | p{0.1\textwidth} | p{0.1\textwidth} | p{0.25\textwidth} | p{0.25\textwidth}}
	
	\textbf{Ascii} & \textbf{Unicode}  \\ \hline
	<==> & u21d4  \\  
	==> & u21d2  \\  
	<== & u21d0  \\  
	\&\& & u2227  \\  
	| | & u2228  \\  
	!= & u2260  \\  
	<= & u2264  \\  
	>= & u2265  \\  
	:: & u2022  \\  
	! & u00ac  \\  
	forall & u2200  \\  
	exists & u2203  
	
	
\end{longtable}

Unfortunately they are not working right know because of a bug in the parser.






