\subsection{Evaluation}
Following the comparison made in \ref{golrech}, it can safely be said that product developed during this project is at least en pair with most existing solutions. An automated setup of the entire environment which resolves dependencies on all major platforms guarantees that new users can make their first steps with Dafny without having to spend time to get things running. This is an important first stepping stone in making Dafny accessible for a wide user base. \newline

After having gotten Dafny to run on an environment, the next concern is to offer a productive, efficient and enjoyable programming experience. This keeps programmers using Dafny interested and keen on further using it. Having implemented many common IDE features as well as some Dafny specific functionality lets programmers apply work flows they have established with other languages as well as discovering the big benefits regarding functional correctness Dafny can offer. \newline

Structuring the plugin as a language server decouples it from any particular IDE, although this project relied on Visual Studio Code as a reference IDE. This decoupling leads to the possibility to integrate Dafny into many other IDEs without having to duplicate any logic. This further lays the groundwork to support a wide user base as a programmer would not have to switch the IDE he has come used to for another one. \newline

It was a major concern for this project to not just develop a proof of concept, but a production ready release of the plugin. Through the application of continuous integration feedback from users could be integrated early on. The plugin has been already downloaded over 300 times from all over the world at the time of this writing. It was also used in a course concentrating on software engineering at a school for higher education to get students familiar with contract based reasoning. The minor changes that had to be made to the Dafny pipeline itself during this project have been accepted by Microsoft and are integrated into the official version of the pipeline. \newline

The open source nature of this project encourages people to participate in enriching the plugin, as well as integrating it in additional IDEs. Close detail has been paid on keeping the documentation as well as the code quality at a high standard, so contributing should be possible without having to invest too much time to get started. It stands to hope that providing the plugin in the state it is at the end of this project is a big stepping stone in widening the user base that can profit from using Dafny. 