\subsection{Main Goal}\label{mainGoal}

The main goal of this project is to make Dafny accessible for a wider user base. In order to this, the focus was laid on two main objectives. \newline

The first one is to provide a simple setup for all the tooling that is necessary to program with Dafny. This includes, next to the IDE itself, also the compiler and the proof engine pipeline which Dafny uses, among other things. The central point of handling the setup does not only make the life of the programmer easier, but also allows for the control of upgrades and dependencies in a uniform manner. \newline

The second objective is to support the programmer when actually writing code. Programmers are used to get rich support from modern IDEs and the question of choosing of a programming language for a project is often interwoven with the quality of tooling behind that language. To make Dafny attractive for a broader range of users, the IDE should offer Dafny specific feature such as help with writing specification constructs. Through this, programmers are best able to learn quickly which features Dafny offers and to use them. Next to the language specific features, standard IDE features which programmers have grown accustomed to must not be forgotten. If a new IDE does not offer support for features such as go to definition, find references, and a minimal set of refactorings, many programmers may stop using it after a short amount of time because the IDE does not support the work flow they have gotten used to. \newline

The main goal of this project therefor lies in providing robust, well working solutions regarding these two objectives. \newline