\section{Management Summary}
Following the quickly growing digitalization of businesses and the therefor more complex applications being developed, two key points are gaining focus in today's IT landscape. The first one is the proven functional correctness of programs and the second one is the uprising of multi threaded applications. Writing multi threaded applications becomes much simples once the functional correctness is proven, thus it can be stated that proven functional correctness is a stepping stone towards the easier implementation of parallelism. Dafny is a programming language which tries to move the focus of writing correct code towards writing correct specification constructs, which is often easier. Applying this concept consistently should result in being able to turn business requirements into correct working implementations quicker as with traditional languages. The proven correlation between the time of discovering an error and the cost of fixing it also strongly advises writing software which is proven correctly as early in the life cycle of the product as possible. Even though using Dafny in business is compelling for these reasons, its usage is still not widespread, something which this project tries to change.\newline

The wide spread usage of a tool for programmers is mainly dictated by two factors, namely the burden of getting it to run and the support that it is able to offer the programmer. \newline

To address the first point, the plugin was developed for Visual Studio Code, an IDE running on all major platforms. It was also given an installation routine which resolves all dependencies on all platforms automatically after pressing one button. This also allows programmers that are not that familiar with the console to rapidly develop Dafny programs, something that was not possible given the tooling existing up until now. \newline

Regarding the second point, it was always paramount to offer as much help as possible to the programmer. This was done in two steps, the first one being implementing standard features that a programmer is used to when working with an IDE, which could be achieved within this project. The second step are language specific features. To offer these, some research was done in what situations often arise when programming Dafny in order to reveal which features a programmer could most benefit from without solving all specification construct suggestions in a generic way. This pragmatic approach helped the project stay in scope while still offering rich help in many common programming contexts. \newline

While at first not a key concern, it was decided to implement the plugin according to an emerging standard in semantic language analysis platforms. This means that the plugin does not only work well with Visual Studio Code, but can be integrated into many other IDEs such as Emacs with very little adjustments, further broadening the possible Dafny user base. \newline

During the project, production quality was always striven for, so that the end product was not a prototype which no one uses. Due to the application of continuous integration the user base of the plugin has already reached 300 people at the time of this writing, proving the usability and robustness of the product developed by this project. The project remains open source, inviting other to continue the work and share the benefits of Dafny with even more people. \newline