\section{Personal Summaries}

\subsection{Personal Summary Markus Schaden}
At the beginning of this project, I was exited but also a little bit nervous about the following weeks. I already did some IDE integration during the term thesis, but the required core features were distinct from the bachelor thesis. Not only did we have to implement a complete new language into Visual Studio Code, but Dafny also is specification construct driven, a programming language type which I rarely worked with. \newline
After the first weeks, we did some good progress, although our project plan was quite ambitious. Especially the preparation looking at the tutorials helped. Unfortunately there was no documentation about the source code. Thereby the only chance to develop new features, was to debug through the code. This helped to gain a understanding about how the verification is working under the hood. The first highlight was, when the export of the symbol table worked for the first time and after a lot of trial and error the project finally successfully compiled.  \newline
During the next couple of weeks my understanding of the Dafny pipeline got better and better. It also became much clearer how the three components are working together. Especially when I started implementing the counter examples, I already knew where to start in order to implement this important feature.  \newline
Due to various other project were I worked with Rafael, I knew that the collaboration with him will be pleasant. During tedious work or if one of us was struggling, it was helpful to work together to overcome problems. Also the discussions and code reviews helped to finish this project in time and with success.   \newline
Farhad Mehta looked after us in a pleasant way. We had the freedom to make decisions and develop, but we also could rely on him if we had problems or needed a third opinion about a topic. Especially he was very helpful in making us understand the proof theory much better. Without this, implementing the counter examples was not possible and we would still be struggling with contract generation, which is not possible with our current knowledge. Also his extensive feedbacks helped to optimize the report and to make the plugin better.  \newline
I really liked to work on this bachelor thesis and am happy with the result. Next to that I learned a lot in various directions, it was a very exiting project. Since we already released the plugin after some weeks, it is a good feeling to know that it is used around the globe. Due to the open source license and the detailed documentation, it would be very nice if some people would start to extend the functionality of the plugin.   \newline
\clearpage
\subsection{Personal Summary Rafael Krucker}
The work on this project has been a very existing and informative experience. While having previously worked on IDE plugins, it was something entirely different integrating an entire language, especially one that is specification constructs driven, something which I also was not very experienced in. \newline
Our project plan was pretty ambitious, while having no experience in proof theory we wanted to integrate a generic way to propose specification. While this task would be very interesting, it soon became apparent that it was not feasible with our knowledge in proof theory and the project scope. Additionally, existing literature suggests that this problem is almost unsolvable hard anyway when working with the kind of proof engine that the Dafny pipeline uses. However, it was very interesting to gain a working knowledge in this area. \newline
After the above mentioned problem was discovered, we reduced our ambitions to provide solutions to concrete problems, which we discovered emerge quite often during programming. It was interesting to isolate situations that programmers could be helped with while it still had to be feasible to provide this help. At the end of this project, I am quite content with the Dafny specific features that we were able to implement, ranging from automatic null checking up unto providing decrease guards. This relieves the programmer of much tedious, but important work. \newline
Farhad Mehta was always quick to break down problems to an understandable point of abstraction when we got stuck somewhere and also provided many important suggestions and ideas throughout our work. This made us progress smoothly while still leaving us room to try to tackle problems we deemed difficult. It is safe to say our plugin would not be as coherent as it is now would it not have been for the iterative feedback loop that he provided very professionally \newline
Next to implementing Dafny specific features, we also implemented all the important IDE agnostic features one expects from a language integration. This was interesting to do and further deepened my knowledge about typescript. The plugin also raised questions about efficiency, correct scoping, caching and the correct architecture which I haven't encountered in this way before and I am proud of the way we resolved them in our product. The implementation of the language server protocol also made our plugin mostly IDE agnostic, something which I also found pretty cool. \newline 
Having already worked on previous projects with Markus, I knew that our collaboration would run smoothly as ever while being very efficient and fruitful. Providing each other with feedback consistently throughout the project helped it reach the quality it has now. Having a similar working speed, we were able to reach our goals and finish the project on time. I would be delighted if all my further collaborations could be held to this standard. \newline
Because we applied the concept of continuous delivery, the plugin already is used by several people around the globe, giving it at least some creditability regarding stability and helpfulness. The documentation and the open source nature of the project invite other people to continue the work on this project and I am happy when it is of use to the Dafny community. 
